 

### 2023-08-15 markdown/article.Rmd
suggestions from supervisor implemented in article
article shortened significantly

ready to implement beamer blocks

\newpage
# Introduction
The central dogma of molecular biology serves as the cornerstone of biological processes, providing a framework for understanding how genetic information is converted into functional proteins \citep{Cobb2017}. Over the past six decades, researchers have explored various omics fields, such as genomics, transcriptomics, proteomics, and metabolomics, to comprehensively analyze biomolecules in diverse contexts. Omics research offers a holistic view of biological processes and regulatory mechanisms, and identify key players in complex biological networks using computational methods. Proteins have various functions, such as providing structural integrity, catalyzing chemical reactions, and regulating cellular functions \citep{Karahalil2016}. Proteomics is a comprehensive approach that investigates the complete protein composition of a specimen, aiming to understand the intricate biological network it represents.

@@

As seen in figure \@ref(fig:ui-numbers-default) the aspects shown in the main panel include:
\spacing{1}
\begin{scriptsize}
\begin{itemize}
\item \textbf{Overview:} Summary showing the individual processing steps of the analysis pipeline
\item \textbf{Summary Barplot:} Barchart showing the number of peptide spectrum matches, peptides and proteins
\item \textbf{Normalization validation:} Mutual information intra- and inter- groupwise
\item \textbf{Reporter ion intensity:} Includes factor selection to create boxplots indicating a factor
\item \textbf{Median coefficient of variation and correlation:} Variation of peptides for razor proteins and correlation heatmap for protein expression
\item \textbf{Dimensionality reduction:} Analysis with principle component analysis (PCA) and Uniform Manifold Approximation and Projection (UMAP)
\item \textbf{Feature wise output:} Visualization of an individual protein over the course of the pipeline.
\item \textbf{Statistics module:} Differential expression analysis
\item \textbf{Pathway enrichment:} Over-representation analysis for differentially expressed proteins in the context of pathways.
\item \textbf{Protein set ontology enrichment:}  Over-representation analysis for differentially expressed proteins in a user selected context.
\end{itemize}
\end{scriptsize}
\spacing{1.5}

\newpage
```{r ui-mutual-info-quant-norm, echo=F, out.width="100%", fig.align='center', fig.cap="Mutual information with quantile normalization"}
knitr::include_graphics("screenshots/mutual_info_quant_norm.png")
```

@@

After performing a logarithmic transformation to the base 2 and quantile normalization, the mutual information (MI) within the healthy control (HC) group's sample type exhibits a minor reduction in the interquartile range (IQR) (as seen in figure \@ref(fig:ui-mutual-info-quant-norm)). However, it should be noted that there is an increase in the number of outliers compared to the raw data.
When switching to the quantile normalization the loss of mutual information between sample types decreased (as seen in figure \@ref(fig:ui-mutual-info-quant-norm)). Therefore the advised normalization technique for this particular data set would be the quantile normalization compared to the default method.

\newpage
```{r ui-stat-HC-HIV-noMets, echo=F, out.width="70%", fig.align='center', fig.cap="Statistics module in portrait mode"}
knitr::include_graphics("screenshots/stat_HC_HIV_noMets.png")
```

\newpage
By comparing the log10-transformed and quantile-normalized expression values between the HIV group without metabolic syndrome and the healthy control (HC) group (see figure \@ref(fig:ui-stat-HC-HIV-noMets)), six differentially expressed proteins were identified. The Venn diagram indicates significantly up/down regulated proteins. The Volcano represents -log10(adjusted p values) against log2 fold changes (logFC).
These proteins include P68133 (Actin, alpha skeletal muscle), P47914 (60S ribosomal protein L29), P25311 (Zinc-alpha-2-glycoprotein), O60506 (Heterogeneous nuclear ribonucleoprotein Q), P31153 (S-adenosylmethionine synthase isoform type-2), and P46777 (60S ribosomal protein L5). The remaining 652 proteins have been identified as not significantly up/down regulated in the comparison analysis. These proteins did not exhibit statistically significant differences in expression between the compared groups or conditions.

@@

\newpage
\listoffigures

>>>>

\newpage
# Introduction
The central dogma underpins genetic information's transformation into functional proteins \citep{Cobb2017}. Proteins have various functions, such as providing structural integrity, catalyzing chemical reactions, and regulating cellular functions \citep{Karahalil2016}. Omics fields like genomics, transcriptomics, proteomics, and metabolomics provide an holistic  view of biomolecules. These proteins are comprehensively studied through proteomics to unveil intricate biological networks.

@@

<!-- As seen in figure \@ref(fig:ui-numbers-default) the aspects shown in the main panel include: -->
<!-- \spacing{1} -->
<!-- \begin{scriptsize} -->
<!-- \begin{itemize} -->
<!-- \item \textbf{Overview:} Summary showing the individual processing steps of the analysis pipeline -->
<!-- \item \textbf{Summary Barplot:} Barchart showing the number of peptide spectrum matches, peptides and proteins -->
<!-- \item \textbf{Normalization validation:} Mutual information intra- and inter- groupwise -->
<!-- \item \textbf{Reporter ion intensity:} Includes factor selection to create boxplots indicating a factor -->
<!-- \item \textbf{Median coefficient of variation and correlation:} Variation of peptides for razor proteins and correlation heatmap for protein expression -->
<!-- \item \textbf{Dimensionality reduction:} Analysis with principle component analysis (PCA) and Uniform Manifold Approximation and Projection (UMAP) -->
<!-- \item \textbf{Feature wise output:} Visualization of an individual protein over the course of the pipeline. -->
<!-- \item \textbf{Statistics module:} Differential expression analysis -->
<!-- \item \textbf{Pathway enrichment:} Over-representation analysis for differentially expressed proteins in the context of pathways. -->
<!-- \item \textbf{Protein set ontology enrichment:}  Over-representation analysis for differentially expressed proteins in a user selected context. -->
<!-- \end{itemize} -->
<!-- \end{scriptsize} -->
<!-- \spacing{1.5} -->

<!-- \newpage -->
<!-- ```{r ui-mutual-info-quant-norm, echo=F, out.width="100%", fig.align='center', fig.cap="Mutual information with quantile normalization"} -->
<!-- knitr::include_graphics("screenshots/mutual_info_quant_norm.png") -->
<!-- ``` -->

@@

After performing a logarithmic transformation to the base 2 and quantile normalization, the mutual information (MI) within the healthy control (HC) group's sample type exhibits a minor reduction in the interquartile range (IQR). However, it should be noted that there is an increase in the number of outliers compared to the raw data.
When switching to the quantile normalization the loss of mutual information between sample types decreased. Therefore the advised normalization technique for this particular data set would be the quantile normalization compared to the default method.

```{r ui-stat-HC-HIV-noMets, echo=F, out.width="60%", fig.align='center', fig.cap="Statistics module in portrait mode"}
knitr::include_graphics("screenshots/stat_HC_HIV_noMets.png")
```

By comparing the log10-transformed and quantile-normalized expression values between the HIV group without metabolic syndrome and the healthy control (HC) group (see figure \@ref(fig:ui-stat-HC-HIV-noMets)), six differentially expressed proteins were identified. The Venn diagram indicates significantly up/down regulated proteins. The Volcano represents -log10(adjusted p values) against log2 fold changes (logFC).
<!-- These proteins include P68133 (Actin, alpha skeletal muscle), P47914 (60S ribosomal protein L29), P25311 (Zinc-alpha-2-glycoprotein), O60506 (Heterogeneous nuclear ribonucleoprotein Q), P31153 (S-adenosylmethionine synthase isoform type-2), and P46777 (60S ribosomal protein L5). The remaining 652 proteins have been identified as not significantly up/down regulated in the comparison analysis. These proteins did not exhibit statistically significant differences in expression between the compared groups or conditions. -->

@@

<!-- \newpage -->
<!-- \listoffigures -->
